Indirect elicitation as a necessary condition for consistent surrogate losses

### 2020-06-02 neurips-20/cvx-flats.tex
Rephrasing beginning of sec 4

Instead of a target loss $\ell$, one may want to learn a conditional statistic such as the expected value, variance, or entropy.
In this case, following the tradition in the statistics literature on conditional estimation
\raft{Jessie: could you add the refs commented out here, and then kill the comment?},
% 1. La ́szlo ́ Gyo ̈rfi Michael Kohler Adam Krzyz ̇ak Harro Walk
% A Distribution-Free Theory of Nonparametric Regression
% 2. Local Polynomial Variance-Function Estimation
% David Ruppert , M. P. Wand , Ulla Holst & Ola HöSJER
% To cite this article: David Ruppert , M. P. Wand , Ulla Holst & Ola HöSJER (1997) Local
% Polynomial Variance-Function Estimation, Technometrics, 39:3, 262-273
% To link to this article: http://dx.doi.org/10.1080/00401706.1997.10485117
% 3. https://academic.oup.com/biomet/article-abstract/85/3/645/229122
we formalize consistency as converging to the correct conditional estimates of the property.
Convergence is measured by functions $\propdis(r, p)$ that formalize how close $r$ is to ``correct'' for conditional distribution $p$.
In particular we should have $\propdis(r,p) = 0 \iff r \in \gamma(p)$.

@@

If $\gamma$ is a property elicited by a target loss $\ell$, then taking $\propdis$ to be the regret with respect to $\ell$, the two definitions are equivalent. % for the appropriate choice of $\propdis$.
% when using target regret of $r$ (compared to an optimal prediction for $p$) as $\propdis(r,p)$.
\begin{lemma}\label{lem:consistent-loss-implies-prop}
	Given a surrogate loss $L$, link $\psi$, and target loss $\ell$, set
	% Take $\propdis: \R \times \simplex \to \reals_+$ with 
  $\mu(r,p) := R_\ell(r,p)$.
  Then
	$(L, \psi)$ is consistent with respect to $\ell$ if and only if $(L,\psi)$ is $\propdis$-consistent with respect to $\gamma := \prop{\ell}$.
\end{lemma}
\begin{proof}
	First, observe that $\propdis(r,p) = 0 \iff r \in \gamma(p)$, since $r \in \gamma(p)$ implies $\ell(r;p) = \risk{\ell}(p)$, and if $r \not \in \gamma(p)$, then $\ell(r;p) > \risk{\ell}(p)$.
  \raft{We don't define $\risk{\ell}$}
	Now suppose $(L, \psi)$ are consistent with respect to $\ell$, and take any sequence $\{f_m\}$ of measurable hypotheses.
  Then
  \raft{Discuss proof}
        \begin{align}
	  &\; \E_D \ell(\psi \circ f_m(X), Y)\to \inf_f \E_D \ell(\psi \circ f(X), Y)   \label{eqn:cons-loss-cond} \\
	  &\iff \E_X R_\ell(\psi \circ f_m(X), D_X) \to 0                               \nonumber  \\
	  &\iff \E_X \propdis(\psi \circ f_m(X), D_X) \to 0~.~                          \label{eqn:cons-prop-cond}
	\end{align}
        Therefore, we have that $\mathbb{E}_D L(f_m(X),Y) \to \inf_f \mathbb{E}_D L(f(X),y)$ implies (\ref{eqn:cons-loss-cond}) if and only if it implies (\ref{eqn:cons-prop-cond}).
\end{proof}

@@

\bo{I like all this, but maybe it can be reworded to flow better with the above?}
Flats of a convex (expected) loss are complemented by its subgradients.\footnote{We define the subdifferential $\partial L(u,y) = \{g \in \reals^d : L(u',y) \geq L(u,y) +g^T (u' - u) \; \forall u' \in \reals^d  \}$, and members of this set subgradients.}  % of convex losses and their flats in order to understand convex elicitation complexity. 
While subgradients live in the prediction space and flats in probability space, the two can be connected considering a $p \in \simplex$.
In a loose sense, this $p$ is a witness of a flat of the level set, and the flat must be of certain codimension (dependent on prediction dimension) so that, for this witness $p$, we can extract a witness $W$ of subgradient sets proving that $\vec 0 \in \partial \exploss{L}{u}{p}$; our desired optimality condition.\jessiet{This is probably TMI but it at least helped me to write it out and think of it this way (p as a witness of the flat, in a loose sense)}
The intuition for this result comes from the proof of~\citet[Theorem 16]{ramaswamy2016convex}, although indirect elicitation lets us simplify their proof that originally relied on calibration.
In fact, their main proof tool, the subspace of feasible dimensions (Definition~\ref{def:subspace-feas}), is a linear shift of a flat satisfying the requirements of Theorem~\ref{thm:cvx-flats}, which lets us deduce their result as Corollary~\ref{cor:fsd-bound} through a much shorter proof thanks to our reliance on elicitation.

>>>>

Instead of a target loss $\ell$, one may want to learn a conditional statistic such as the expected value, variance, or entropy.
In this case, following the tradition in the statistics literature on conditional estimation~\citep{gyorfi2006distribution,fan1998efficient,ruppert1997local},
we formalize consistency as converging to the correct conditional estimates of the property.
Convergence is measured by functions $\propdis(r, p)$ that formalize how close $r$ is to ``correct'' for conditional distribution $p$.
In particular we should have $\propdis(r,p) = 0 \iff r \in \gamma(p)$.

@@

If $\gamma$ is a property elicited by a target loss $\ell$, then taking $\propdis$ to be the regret with respect to $\ell$, the two definitions are equivalent. % for the appropriate choice of $\propdis$.
% when using target regret of $r$ (compared to an optimal prediction for $p$) as $\propdis(r,p)$.
\begin{lemma}\label{lem:consistent-loss-implies-prop}
	Given a surrogate loss $L$, link $\psi$, and target loss $\ell$, set
	% Take $\propdis: \R \times \simplex \to \reals_+$ with 
  $\mu(r,p) := R_\ell(r,p)$.
  Then
	$(L, \psi)$ is consistent with respect to $\ell$ if and only if $(L,\psi)$ is $\propdis$-consistent with respect to $\gamma := \prop{\ell}$.
\end{lemma}
\begin{proof}
	First, observe that $\propdis(r,p) = 0 \iff r \in \gamma(p)$, since $r \in \gamma(p)$ implies $\ell(r;p) = \risk{\ell}(p) := \inf_{r' \in \R} \exploss{\ell}{r'}{p}$, and if $r \not \in \gamma(p)$, then $\ell(r;p) > \risk{\ell}(p)$.
  \raft{We don't define $\risk{\ell}$}\jessiet{Added in line, but very subtle.  not a huge fan.}
	Now suppose $(L, \psi)$ are consistent with respect to $\ell$, and take any sequence $\{f_m\}$ of measurable hypotheses.
  Then
  \raft{Discuss proof}
        \begin{align}
	  &\; \E_D \ell(\psi \circ f_m(X), Y)\to \inf_f \E_D \ell(\psi \circ f(X), Y)   \label{eqn:cons-loss-cond} \\
	  &\iff \E_X R_\ell(\psi \circ f_m(X), D_X) \to 0                               \nonumber  \\
	  &\iff \E_X \propdis(\psi \circ f_m(X), D_X) \to 0~.~                          \label{eqn:cons-prop-cond}
	\end{align}
        Therefore, we have that $\mathbb{E}_D L(f_m(X),Y) \to \inf_f \mathbb{E}_D L(f(X),y)$ implies (\ref{eqn:cons-loss-cond}) if and only if it implies (\ref{eqn:cons-prop-cond}).
\end{proof}

@@

\bo{I like all this, but maybe it can be reworded to flow better with the above?}
Flats of a convex (expected) loss are complemented by its subgradients.\footnote{We define the subdifferential $\partial L(u,y) = \{g \in \reals^d : L(u',y) \geq L(u,y) +g^T (u' - u) \; \forall u' \in \reals^d  \}$, and members of this set subgradients.}  % of convex losses and their flats in order to understand convex elicitation complexity.
Fixing a distribution $p \in \simplex$, we observe the reports $u$ where $\vec 0$ is in the $p$-weighted Minkowski sum of the subdifferentials $\partial L(u,y)$, where the summation is over $y$.
By definition of the Minkowski sum, this yields a vector of witnesses that are subgradients whose $p$-weighted sum is exactly $\vec 0$.
Moreover, this exact choice of subgradients must add to $\vec 0$ for all $q$ in the same flat containing $p$, and the dimension of the flat is determined in part by the prediction dimension.
%While subgradients live in the prediction space and flats in probability space, the two can be connected considering a $p \in \simplex$.
%In a loose sense, this $p$ is a witness of a flat of the level set, and the flat must be of certain codimension (dependent on prediction dimension) so that, for this witness $p$, we can extract a witness $W$ of subgradient sets proving that $\vec 0 \in \partial \exploss{L}{u}{p}$; our desired optimality condition.\jessiet{This is probably TMI but it at least helped me to write it out and think of it this way (p as a witness of the flat, in a loose sense)}
The intuition for this result comes from the proof of~\citet[Theorem 16]{ramaswamy2016convex}, although indirect elicitation lets us simplify their proof that originally relied on calibration.
In fact, their main proof tool, the subspace of feasible dimensions (Definition~\ref{def:subspace-feas}), is a linear shift of a flat satisfying the requirements of Theorem~\ref{thm:cvx-flats}, which lets us deduce their result as Corollary~\ref{cor:fsd-bound} through a much shorter proof thanks to our reliance on elicitation.
